---
title: "Tissue loss vs. temperature plots"
author: "Julianna Renzi"
date: "11/7/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clear the workspace
rm(list = ls())

# load required packages
require(tidyverse) # data manipulation
require(here) # relative file paths
require(lubridate) # dates and times
require(wesanderson) # color palette
require(patchwork) # combining figures
```

# Bring in the data

```{r}
# this is the temperature time series from the experiment, with one value per date/time (averaged across HOBO loggers, with a value per few minutes)
tempTimeseries <- read_csv(here("july2024_updates/generated_data/hobo_timeseries.csv")) # this file was generated using the temperature_wrangling.Rmd script

# this is the percent tissue loss change experienced by a coral on a given day 
tissueLoss <- read_csv(here("july2024_updates/generated_data/tissue_loss_change.csv")) # this file was generated using the daily_tissue_loss_change_wrangling.Rmd script
```

# Set aesthetics

Create our color palette from the Wes Anderson package

```{r}
vibRed <- wes_palette("Darjeeling1", n = 5)[1]
forGreen <- wes_palette("Darjeeling1", n = 5)[2]
lightO <- wes_palette("Darjeeling1", n = 5)[3]
darkO <- wes_palette("Darjeeling1", n = 5)[4]
skyB <- wes_palette("Darjeeling1", n = 5)[5]


# and specify text size
txt_sz <- 15 
```


# Plot temperature through time

Look at how the temperature changes over time

```{r}
tempTimeseries %>% 
  filter(Date_time > as_datetime("2020-02-01 00:00:00")) %>% # just from the start of the tissue loss data
  ggplot(aes(x = Date_time, y = Mean_temp)) +
  geom_line() +
  geom_hline(yintercept = 27.3, 
             color = vibRed, 
             linetype = "dotted") + # this is the Heron monthly mean water temp according to: https://seatemperature.info/heron-island-water-temperature.html and also JB
  theme_classic() +
  # make first rectangle highlighting the first peak 
  annotate("rect",
           xmin = as_datetime("2020-02-03 00:00:00"), 
           xmax = as_datetime("2020-02-08 00:00:00"), 
           ymin = -Inf, 
           ymax = Inf, fill = lightO, alpha = 0.2) +
  annotate("rect",
           xmin = as_datetime("2020-02-15 00:00:00"), 
           xmax = as_datetime("2020-02-22 00:00:00"), 
           ymin = -Inf, 
           ymax = Inf, fill = lightO, alpha = 0.2) +
  ylab("Water temperature (°C)") +
  xlab("") +
  ylim(26, 33)  +
  # change text size
  theme(text = element_text(size = txt_sz)) +
  ggtitle("(a)") -> tempTSFig
  

```


# Plot tissue loss through time

Get the average tissue loss by treatment (average tissue loss across the treatment group on a given day)

```{r}
tissueLoss %>% 
  group_by(Crab_treatment, Algae_treatment, Wounding_treatment, Date) %>%
  summarize(MeanTL = mean(Perc_tissue_loss)) -> avgTL
```

Then plot the data through time, broken up by treatments

```{r}
# make labels nice
crab.labs <- c("Crab", "No crab")
names(crab.labs) <- c("Y", "N")
```

First for crab:

```{r}
avgTL %>%
  ggplot(aes(x = Date, y = MeanTL, color = Algae_treatment, linetype = Wounding_treatment)) +
  geom_line() +
  scale_linetype_manual(values=c("solid", "dashed"),
                        name = "Wounding") +
  facet_grid(.~Crab_treatment, labeller = labeller(Crab_treatment = crab.labs)) + # facet by crabs
  ylab("Average tissue loss (%)") +
  xlab("") +
  scale_color_manual(values = c(darkO, forGreen),
                     name = "Algae") + # clean up legend and colors
  theme_classic() +
  # add rectangles showing extreme heat periods
  annotate("rect",
           xmin = as.Date("2020-02-03", "%Y-%m-%d"), 
           xmax = as.Date("2020-02-08", "%Y-%m-%d"), 
           ymin = -Inf, 
           ymax = Inf, fill = lightO, alpha = 0.2) +
  annotate("rect",
           xmin = as.Date("2020-02-15 00:00:00", "%Y-%m-%d"), 
           xmax = as.Date("2020-02-22 00:00:00", "%Y-%m-%d"), 
           ymin = -Inf, 
           ymax = Inf, fill = lightO, alpha = 0.2) +
  # change text size
  theme(text = element_text(size = txt_sz),
        legend.position = "top") +
    ylim(0, 43) +
  ggtitle("(b)") -> TLfig
```





# Combine and save figure 

Save these figures as a PDF in the figures folder

```{r}
pdf(file = here("july2024_updates/figures/tissue_loss_temperature.pdf"), width = 8.5, height = 9)

tempTSFig / TLfig

dev.off()
```

