---
title: "Crab feeding behavior"
author: "Julianna Renzi"
date: "4/21/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clear the workspace
rm(list = ls())

# load required packages
require(here) # relative file paths
require(tidyverse) # data wrangling
require(RColorBrewer) # color palette
require(FSA) # for dunn test
```

# Bring in the data

```{r load data}
# bring in the csv that has the percent cover of each nubbin, determined beforehand using ImageJ
perCover <- read_csv(here("data/crab_feeding_imageJ.csv"))

# bring in the bite count data from the videos
behavior <- read_csv(here("data/crab_feeding_behavior.csv"))
```

# Get carapace information

To report in text

```{r carapace size}
mean(behavior$Carapace_mm)

# standard error
sd(behavior$Carapace_mm)/sqrt(length(behavior$Carapace_mm))
```

# Calculate Strauss' index to assess preference

Join the dataframes by the trial number and add derived columns (e.g., total number of bites, percent cover, Strauss' index, etc.). Strauss' linear selectivity index was taken from the Transactions of the American Fisheries Society, 1979 ("Reliability estimates fro Ivlev's electivity index, the forage ratio, and a proposed linear index of food selection") and is just the difference in the proportion of bites taken vs. the availability of the resource. That means that the index ranges from -1 to 1, with positive values indicating a preference, negative values indicating avoidance, and 0 indicating no preference. It also means that t-statistics can be used for statistical comparison, according to Strauss.

```{r calculate Strauss Index}
behavior %>% 
  full_join(perCover, by = "Trial") %>% 
  # add columns for percent cover of relevant pixels (alive/margin/dead-where we counted bites)
  mutate(Total_pixels = Alive_Pixels + Margin_Pixels + Dead_Pixels,
         Prop_dead = Dead_Pixels / Total_pixels,
         Prop_margin = Margin_Pixels / Total_pixels,
         Prop_alive = Alive_Pixels / Total_pixels) %>% 
  # add columns for percent of bites
  mutate(Total_bites = Bites_alive + Bites_margin + Bites_dead,
         Prop_bites_dead = Bites_dead / Total_bites,
         Prop_bites_margin = Bites_margin / Total_bites,
         Prop_bites_alive = Bites_alive / Total_bites) %>% 
  # calculate the indices
  mutate(Dead = Prop_bites_dead - Prop_dead,
         Margin = Prop_bites_margin - Prop_margin,
         Alive = Prop_bites_alive - Prop_alive) %>% 
  # remove the crabs that did not feed
  filter(Total_bites != 0) -> strauss
```

Also reformat for a few tasks

```{r get data in long format}
strauss %>% 
  # reformat so it's easier to plot
  pivot_longer(cols = c(Dead, Margin, Alive), names_to = "Substrate", values_to = "Preference") -> strauss_long
```

# Test differences

Are these values different than zero?

```{r}
t.test(strauss$Alive, mu = 0) # is different
  shapiro.test(strauss$Alive) # is normal
  
t.test(strauss$Dead, mu = 0) # is different
  shapiro.test(strauss$Dead) # is normal
  
t.test(strauss$Margin, mu = 0) # is different
  shapiro.test(strauss$Margin) # is normal


# these are signifiant even with a Bonferonni correction for doing 3 separate tests:
0.05/3
```

Are these values different from each other? 

```{r}
kruskal.test(Preference ~ Substrate, data = strauss_long) # this is non-parametric

m1 <- aov(Preference ~ Substrate, data = strauss_long) # this is parametric
  summary(m1) # they are different either way


dunnTest(Preference ~ as.factor(Substrate),
         data = strauss_long,
         method="bonferroni")
```

# Plot 

```{r plot Strauss index vs. treatment}
# want to order it as alive, margin, dead for aesthetics
level_order <- c('Alive', 'Margin', 'Dead') 

pdf(here("figures/fig_S4_feeding_pref.pdf"), width = 6, height = 4)

strauss_long %>% 
  # plot
  ggplot(aes(x = factor(Substrate, level = level_order), y = Preference, fill = Substrate)) +
  geom_boxplot() +
  labs(x = "Substrate", y = "Preference") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  scale_fill_brewer(palette="Oranges") + 
  geom_point(alpha = 0.5) +
  theme_classic() +
  theme(text = element_text(size=18), legend.position="none") +
  xlab("") +
  ylim(-1, 1) +
  stat_summary(geom = 'text', label = c("a", "b", "c"), fun.y = max, vjust = -1, size=6, color = c("black", "black", "black")) # add letters for significance

dev.off()

```





