---
title: "Field data explore"
author: "Julianna Renzi"
date: "3/3/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clear the workspace
rm(list = ls())

# load required packages
require(tidyverse)
require(here)
require(nlme)
require(emmeans)
require(RColorBrewer)
require(survival) # for creating survival objects
require(coxme) # for creating cox mixed effects models
```

# Read in algae data

Note: "NW" is no wounding and "W" is wounding. See methods for more details.

```{r}
# algal weight pre/post experiment
algae <- read_csv(here("data/field_experiment/algae_weight.csv"))

# information on colony size (for summary stat in text)
size <- read_csv(here("data/field_experiment/colony_specs.csv")) 

# daily monitoring of whether algae was present or not
monitoring <- read_csv(here("data/field_experiment/algal_monitoring.csv")) 
```

# Get average colony size 

```{r}
# mean diameter
mean(size$length_cm)
  
# SE diameter
sd(size$length_cm)/sqrt(length(size$length_cm))
```

# Biomass data

## Make sure initial treatments were the same

```{r get initial weight df}
algae %>% 
  select(NW_algae_initial_weight, W_algae_intial_weight, Colony_ID) %>% 
  pivot_longer(cols = c(NW_algae_initial_weight, W_algae_intial_weight),
               names_to = "Treatment", 
               values_to = "Biomass") -> init_biomass

```

```{r calculate mean starting biomass}
init_biomass %>% 
  summarize(Mean_biom = mean(Biomass),
            sq_N = sqrt(n()),
            sd_biom = sd(Biomass),
            se_biom = sd_biom/sq_N)
```

```{r look for differences in starting weights}
init.M <- aov(Biomass ~ Treatment, data = init_biomass)
  summary(init.M) # looks okay (p > 0.05)
```

## Compare treatments

```{r calculate percent change}
algae %>% 
  # get absolute change
  mutate(NW_change = NW_algae_initial_weight - NW_final_weight,
         W_change = W_algae_intial_weight - W_final_weight) %>%
  # get percent change
  mutate(NW_perc_change = NW_change/NW_algae_initial_weight,
         W_perc_change = W_change/W_algae_intial_weight) %>%
  select(Colony_ID, NW_perc_change, W_perc_change) %>% 
  pivot_longer(cols = c(NW_perc_change, W_perc_change),
               names_to = "Treatment", 
               values_to = "Percent_change") -> perc_change
```

## Plot results

```{r plot percent differences}
# pdf(here("figures/field_algal_loss.pdf"), height = 4, width = 6)

perc_change %>% 
  # get this in terms of percents (out of 100)
  mutate(Percent_change = 100*Percent_change) %>% 
  ggplot(aes(x = Treatment, y = Percent_change, fill = Treatment)) +
  geom_boxplot() +
  ylab("Biomass loss (%)") +
  scale_fill_brewer(palette="Oranges") + 
  geom_point(alpha = 0.5) +
  theme_classic() +
  theme(text = element_text(size=18), legend.position="none") +
  xlab("") +
  # rename categories
  scale_x_discrete(labels = c('No wounding','Wounding'))

# dev.off()
```

## Model

We want to model the difference in biomass loss, with a random effect for coral colony, since we use a paired design with 1 patch of each treatment/coral in case there are differences in flow, community composition, etc.

```{r}
m.PercRE <- lme(Percent_change~Treatment, 
                random=~1|Colony_ID, 
                data=perc_change)
  summary(m.PercRE)
  anova(m.PercRE) 
  emmeans(m.PercRE, "Treatment")
```

# Monitoring

Examine changes in clump presence through time using the monitoring data

```{r plot}
# pdf(here("figures/field_monitoring.pdf"), height = 5.5, width = 7)

monitoring %>% 
  # clean up for figure
  mutate(treatment = case_when(treatment == "NW" ~ "No wounding",
                               TRUE ~ "Wounding")) %>% 
  mutate(`Algae present` = case_when(algae_remaining_yn == 1 ~ "Y",
                                     TRUE ~ "N"
                                     )) %>% 
  ggplot(aes(x = time_point, fill = `Algae present`, group = `Algae present`)) +
  geom_bar(stat = "count") +
  facet_wrap(~treatment, ncol = 1) +
  geom_hline(yintercept = 18, color = "gray", linetype = "dashed") +
  geom_hline(yintercept = 16, color = "black", linetype = "dashed") +
  theme_classic() +
  theme(text = element_text(size=18)) +
  xlab("Day") +
  ylab("Number of replicates") +
  scale_fill_manual(values = c(wes_palette("Darjeeling1", n = 5)[4],
                    wes_palette("Darjeeling1", n = 5)[2]))

# dev.off()
```

# Survival analysis

Referee 2 recommended doing a formal analysis here, which we are trying. See below:

```{r make survival dataframe}
# get a key of all patches
monitoring %>% 
  group_by(Colony_ID, treatment) %>% 
  summarize(N = n()) %>% 
  select(-N) -> colonies

# calculate days the algae were lost
monitoring %>% 
  filter(algae_remaining_yn == 0) %>% 
  # get the first day a patch lost it's algae
  group_by(Colony_ID, treatment) %>% 
  summarize(first_day_gone = min(time_point)) %>% 
  full_join(colonies) %>% 
  mutate(censor = case_when(is.na(first_day_gone) ~ 0,
                            TRUE ~ 1)) %>% 
  mutate(first_day_gone = case_when(is.na(first_day_gone) ~ 6,
                                    TRUE ~ first_day_gone)) -> algal_survival
```


```{r make survival object}
surv_object_alg <- Surv(time = algal_survival$first_day_gone, # time to death
                    event = algal_survival$censor) # Censored points: 0 = right censored (not a true death), 1 = event at time
```

Make a Cox model

```{r full model}
cox.model <- coxph(surv_object_alg ~ treatment + cluster(Colony_ID), 
                    data = algal_survival) # note response here must be a survival object returned by the Surv funtion (why we make the surv_object above)
  summary(cox.model)
```

# Test model assumptions

We want to test that the hazards are proportional (don't change over time), which is assumed in the coxph model. We'll use scaled Schoenfeld residuals, which are independent of time and should have a random pattern. The random assumption is supported by p > 0.05

```{r test Schoenfeld}
test.ph <- cox.zph(cox.model) 
  test.ph # meets this assumption (p > 0.05)
  ggcoxzph(test.ph) # no pattern with time
```

Then we'll also test whether there's strong outlier pull/influential observations

```{r look for outliers}
ggcoxdiagnostics(cox.model, type = "dfbeta", # dfbeta plots the estimated changes in the regression coefficients upon deleting each observation in turn
                 ggtheme = theme_bw()) 
```

- These seem all right