---
title: "Examine initial differences in fragment size and any initial tissue loss"
output: html_document
date: "2024-01-05"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clear the workspace
rm(list = ls())

# load required packages
require(tidyverse) # data manipulation
require(here) # relative file paths
require(lubridate) # dates and times
require(wesanderson) # color palette
require(patchwork) # combining figures
require(survival) # survival analysis
require(survminer) # plotting survival curves
```

This script is making sure there are no size differences or differences in initial tissue loss among fragments that are biasing results (e.g., if all the fragments in algal treatments were small, etc.)

# 1. Was initial size important?

## Bring in weight data

```{r}
weights <- read_csv(here("data/initial_buoyant_weights.csv"))

metadat <- read.csv(here("data/metadata.csv")) 
```


## Look at initial differences between treatments

```{r}
weights %>% 
  # join weights with metadata
  full_join(metadat) -> weightCor

# plot this
weightCor %>% 
  ggplot(aes(x = Treatment, y = Initial_dry_weight)) +
  geom_boxplot() # pretty similar across treatments
```

```{r}
mWeight <- aov(Initial_dry_weight ~ Treatment, data = weightCor)
  summary(mWeight) # means appear to be the same
```

## Include in survival analysis

Bring in the survival dataframe:

```{r}
surv_df <- read_csv(here("generated_data/coral_survival37.csv")) %>% # from the survival_wrangling.Rmd script
  # join with initial weights
  full_join(weights)
```

See `cox_model.Rmd` for more commentary on this. Just repeating the analysis here

```{r}
surv_object <- Surv(time = surv_df$Delta_time, # time to death
                    event = surv_df$Censored) # Censored points: 0 = right censored (not a true death), 1 = event at time
```

### Model selection

Make a full model:

```{r}
full.coxphW <- coxph(surv_object ~ Crab_treatment * Algae_treatment * Wounding_treatment + 
                       Initial_dry_weight + cluster(Tank_ID), 
                    data = surv_df) # note response here must be a survival object returned by the Surv funtion (why we make the surv_object above)
```

Winnow down terms:

```{r}
coxph.1W <- coxph(surv_object ~ Crab_treatment + Algae_treatment + Wounding_treatment +
                   Crab_treatment:Algae_treatment +
                   Crab_treatment:Wounding_treatment +
                   Algae_treatment:Wounding_treatment + 
                    Initial_dry_weight + cluster(Tank_ID),
                 data = surv_df)

```

```{r}
coxph.2W <- coxph(surv_object ~ Crab_treatment + Algae_treatment + Wounding_treatment +
                   Crab_treatment:Wounding_treatment +
                   Algae_treatment:Wounding_treatment + 
                    Initial_dry_weight + cluster(Tank_ID),
                 data = surv_df)
```

```{r}
coxph.3W <- coxph(surv_object ~ Crab_treatment + Algae_treatment + Wounding_treatment +
                   Crab_treatment:Wounding_treatment + 
                    Initial_dry_weight + cluster(Tank_ID),
                 data = surv_df)
```

```{r}
coxph.4W <- coxph(surv_object ~ Crab_treatment + Algae_treatment + Wounding_treatment + 
                    Initial_dry_weight + cluster(Tank_ID),
                 data = surv_df)
```

```{r}
coxph.5W <- coxph(surv_object ~ Crab_treatment + Algae_treatment + 
                   Initial_dry_weight + cluster(Tank_ID),
                 data = surv_df)
```

```{r}
coxph.6W <- coxph(surv_object ~ Algae_treatment + 
                   Initial_dry_weight + cluster(Tank_ID),
                 data = surv_df)
```


```{r}
AIC(full.coxphW, coxph.1W, coxph.2W, coxph.3W, coxph.4W, coxph.5W, coxph.6W) # it still looks like the best model is Model 3


# set it as our model
mCoxW <- coxph.3W
  summary(mCoxW) # can see that the effect of size does not seem important
```

Also double check there's no big model performance difference with it in:

```{r}
coxph.3R <- coxph(surv_object ~ Crab_treatment + Algae_treatment + Wounding_treatment +
                   Crab_treatment:Wounding_treatment + cluster(Tank_ID),
                 data = surv_df)

AIC(coxph.3, coxph.3W) # looks like no big difference with the size/weight term
```

# 2. Was initial tissue loss important?

Some corals had small areas of tissue loss at the start of the experiment. We use the code below to explore whether that may have influenced our conclusions.

First, bring in the tissue loss data:

```{r}
# tissue loss estimates
tissueLoss <- read_csv(here("data/coral_tissue_loss.csv"))

# Note: survival data already brought in in part #1 (surv_df)
```

## Initial TL vs. survival

Get a variable for starting tissue loss for each coral

```{r}
tissueLoss %>% 
  # grab data from the start of the experiment
  filter(Date == "1/31/20") %>% 
  # get the average starting tissue loss (mean of both sides of the coral)
  group_by(Coral_ID) %>% 
  summarize(Starting_TL = mean(Percent_dead)) %>% 
  # join with survival data
  full_join(surv_df) -> startTL 
```

Plot

```{r}
startTL %>% 
  ggplot(aes(x = Starting_TL, y = Delta_time, group = Starting_TL)) +
  geom_boxplot() +
  geom_point() +
  theme_bw() # doesn't really look like a trend
```

Is there a difference between starting TL in treatments?

```{r}
mTL <- aov(Starting_TL ~ Treatment, data = startTL)
  summary(mTL) # no
```

## Re-run survival analysis 

Do any differences here influence our conclusions? Re-run analyses:

```{r}
surv_object <- Surv(time = startTL$Delta_time, # time to death
                    event = startTL$Censored) # Censored points: 0 = right censored (not a true death), 1 = event at time
  
```

Create a full cox proportional hazards regression model. Note that a hazard ratio (HR) > 1 is an increased risk of death and HR < 1 is a decreased risk.

```{r}
full.coxphTL <- coxph(surv_object ~ Crab_treatment * Algae_treatment * Wounding_treatment + 
                        Starting_TL + cluster(Coral_ID), 
                    data = startTL) # note response here must be a survival object returned by the Surv funtion (why we make the surv_object above)

```

### Model selection

Now we want to winnow down the terms

```{r}
coxph.1TL <- coxph(surv_object ~ Crab_treatment + Algae_treatment + Wounding_treatment +
                   Crab_treatment:Algae_treatment +
                   Crab_treatment:Wounding_treatment +
                   Algae_treatment:Wounding_treatment + 
                     Starting_TL + cluster(Coral_ID),
                 data = startTL)
```

```{r}
coxph.2TL <- coxph(surv_object ~ Crab_treatment + Algae_treatment + Wounding_treatment +
                   Crab_treatment:Wounding_treatment +
                   Algae_treatment:Wounding_treatment + 
                   Starting_TL + cluster(Coral_ID),
                 data = startTL)
```

```{r}
coxph.3TL <- coxph(surv_object ~ Crab_treatment + Algae_treatment + Wounding_treatment +
                   Crab_treatment:Wounding_treatment + 
                   Starting_TL + cluster(Coral_ID),
                 data = startTL)
```

```{r}
coxph.4TL <- coxph(surv_object ~ Crab_treatment + Algae_treatment + Wounding_treatment + 
                     Starting_TL + cluster(Coral_ID),
                 data = startTL)
```

```{r}
coxph.5TL <- coxph(surv_object ~ Crab_treatment + Algae_treatment + 
                     Starting_TL + cluster(Coral_ID),
                 data = startTL)
```

```{r}
coxph.6TL <- coxph(surv_object ~ Algae_treatment + 
                     Starting_TL + cluster(Coral_ID),
                 data = startTL)
```

Compare models:

```{r}
AIC(full.coxphTL, coxph.1TL, coxph.2TL, coxph.3TL, coxph.4TL, coxph.5TL, coxph.6TL) # It still looks like the best model is Model 3

# check to make sure model results seem the same with startingTL included
summary(coxph.3TL) # see coefficients/significance seem the same and there is no strong effect of starting TL

# Also check whether there's a difference in model fit with or without starting TL
AIC(coxph.3TL, coxph.3) # if anything the one without tissue loss is better
```





