---
title: "Create survival dataframe"
output: html_document
date: "2024-07-06"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clear the workspace
rm(list = ls())

# load packages
require(here) # relative file paths
require(tidyverse)
```

This code creates a survival dataframe from our percent cover estimates for use in survival analyses and figures

# Bring in the data

```{r}
# this is the percent tissue loss on a coral on a given day (cumulative, estimated visually from photos)
coralTL <- read_csv(here("july2024_updates/data/coral_tissue_loss.csv"))
  
# this is the metadata for the experiment
metadat <- read.csv(here("july2024_updates/data/metadata.csv")) 
```

# Calculate percent tissue loss

Since two cover estimates were made on each coral fragment (front and back), we average the two estimates to get the average percent of tissue a coral has lost on a given day

```{r}
# summarize for each date (average both coral sides)
coralTL %>% 
  # get dates in datetime format
  mutate(Date = as.Date(Date, format='%m/%d/%y')) %>% 
  # average for both sides (front & back)
  group_by(Coral_ID, Date) %>% 
  summarize(Mean_TL = mean(Percent_dead)) -> coralMTL # mean tissue loss
```

# Make a survival dataframe 

This code analyses the survival data with the most conservative threshold of 37.5% tissue loss (i.e., the least amount of tissue loss exhibited by a coral fragment that was pulled from the experiment for sampling), which was recommended by a previous reviewer and is likely a better approach (compared to the actual day they were pulled, which could vary in final percent tissue loss), although note that the overall conclusions are the same for both approaches.

```{r}
coralMTL %>% 
  # look for the first day a coral hit this % tissue loss
  filter(Mean_TL >= 37.50) %>% 
  group_by(Coral_ID) %>% 
  summarize(Death_day = min(Date)) %>% 
  # join with metadata
  full_join(metadat) %>% 
  # deal with the two corals that broke--these were not true "deaths" but they were removed from the experiment on these specific days
  mutate(Death_day = case_when(Coral_ID == 7 ~ as.Date("2020-02-14"), 
                               Coral_ID == 29 ~ as.Date("2020-02-16"),
                               TRUE ~ Death_day)) %>% 
  # make sure operations are rowwise (not just the first row)
  rowwise() %>% 
  # since the first day of the experiment was the first day of the month, we can just grab the day of the month as the day of the experiment
  # To do this we turn it into a character to use stringr split functions, but then turn it back into a numeric 
  mutate(Delta_time = str_split(as.character(Death_day), "-")[[1]][3]) %>% 
  mutate(Delta_time = as.numeric(Delta_time)) %>% 
  # censor the ones that broke as well as the ones that made it to the end without >= 37.5% tissue loss (i.e., those with no death days)
  mutate(Censored = case_when(Coral_ID == 7 ~ 0, 
                              Coral_ID == 29 ~ 0,
                              is.na(Death_day) ~ 0,
                              TRUE ~ 1)) %>% 
  # those who stayed in the experiment made it 25 days without "dying", which we'll add in manually here
  # note these aren't true "death days" this is just saying they were censored at day 25 (since we don't know if they would have died on day 26, etc.)
  mutate(Delta_time = case_when(is.na(Death_day) ~ 25,
                               TRUE ~ Delta_time)) %>% 
  mutate(Death_day = case_when(is.na(Death_day) ~ as.Date("2020-02-25"),
                               TRUE ~ Death_day) ) -> surv_df
```

# Write CSV

Save our survival dataframe as a CSV in the `generated_data` folder

```{r}
write_csv(surv_df, here("july2024_updates/generated_data/coral_survival37.csv"))
```

