---
title: "Temperature data wrangling"
author: "Julianna Renzi"
date: "9/28/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clear the workspace
rm(list = ls())

# load required packages
require(tidyverse) # data manipulation
require(here) # relative file paths
require(lubridate) # dates and times
```

# Bring in the temperature data from HOBO temperature loggers inside tanks and their associated metadata

We have a CSV for each logger and we use the code below to combine those into one dataframe with all of the temperature data.

We'll start by bringing in one CSV and then do that same process with a loop for the rest of the CSVs

```{r}
# first get list of all files we need to read in
hob_filelist <- list.files(here("data/HOBO_temperature_loggers/CSVs")) 

# start with the first file on the list
hob_ID <- hob_filelist[1] 
hob_ID <- strsplit(hob_ID, "[.]") # get the file name separated from the .csv extension
hob_ID <- hob_ID[[1]][1] # get logger ID
name <- paste("h", hob_ID, sep="") # get name of the file that we'll use to refer to the csv (h for HOBO)

# rename columns to be more useful
name <- read_csv(here(paste("data/HOBO_temperature_loggers/CSVs/", hob_filelist[1], sep = "")), skip=1)  
  colnames(name)[1] <- "Data_point"
  colnames(name)[2] <- "Date_time"
  colnames(name)[3] <- "Temp_C" 
  colnames(name)[4] <- "Intensity_lum_per_ft2" 
  
# make column into a datetime object to make it easier using lubridate
  # see https://evoldyn.gitlab.io/evomics-2018/ref-sheets/R_lubridate.pdf for more details
name$Date_time <- mdy_hms(name$Date_time)
name <- name[, c(1,2,3)] # remove End of File column/other columns so they don't get wacky in a row bind
name$Hobo_ID <- hob_ID
temp_data <- name

```

Now create the loop

```{r}
# use a loop to import all of the data
for (file in hob_filelist[-1]) { # skip the first file because that's what we did above
  hob_ID <- file
  hob_ID <- strsplit(hob_ID, "[.]")
  hob_ID <- hob_ID[[1]][1]
  name <- paste("h", hob_ID, sep="")
  name <- read_csv(here(paste("data/HOBO_temperature_loggers/CSVs/", file, sep = "")), skip=1) # read in each csv
  colnames(name)[1] <- "Data_point" # re-name columns to match
  colnames(name)[2] <- "Date_time"
  colnames(name)[3] <- "Temp_C"
  colnames(name)[4] <- "Intensity_lum_per_ft2" 
  name$Date_time <- mdy_hms(name$Date_time) # make it a datetime object
  name <- name[, c(1,2,3)] 
  name$Hobo_ID <- hob_ID # add the hobo identifier
  temp_data <- bind_rows(temp_data, name)
}
  
```

Attach the metadata

```{r}
# Bring in the a key for logger tank position and site
hob_metadat <- read_csv(here("data/HOBO_temperature_loggers/HOBO_metadata.csv"))

# want to make key the same format (both characters because tidyverse)
hob_metadat$Hobo_ID <- as.character(hob_metadat$Hobo_ID)

# and check to make sure the IDs match (no obvious typos)
l1 <- sort(unique(as.numeric(temp_data$Hobo_ID)))
l2 <- sort(unique(as.numeric(hob_metadat$Hobo_ID)))

for (id in l1){
  if (id %in% l2){
    print("YEE")
  } else {
    print("Nawhh")
  }
} 

# or:
l1 == l2

# then join the metadata to the temperature data
label_temp <- left_join(temp_data, 
                        hob_metadat, 
                        by="Hobo_ID", 
                        relationship = "many-to-many") 
```

# Begin summarizing the temperature data

```{r}
# look for earliest date
label_temp$Date_time %>% min # Jan 26 (corals acclimating)

label_temp %>% 
  filter(Tank != 21) %>% # don't want the control tank we didn't use
  filter(Date_time >=  as.Date("2020-01-27") & Date_time <  as.Date("2020-02-26")) %>% # just the experiment dates and acclimation period
  mutate(Date_time = ifelse(Hobo_ID == as.character(10550960) & Date_time >  as.Date("2020-02-22"), NA, Date_time)) %>% 
  # get rid of loggers after the dates they were pulled
  # We're only keeping up to the date BEFORE they were pulled so there's no air temperature values from when they were pulled out
  mutate(Date_time = ifelse(Hobo_ID == as.character(10551953) & Date_time >  as.Date("2020-02-05"), NA, Date_time)) %>% 
  mutate(Date_time = ifelse(Hobo_ID == as.character(10551958) & Date_time >  as.Date("2020-02-11"), NA, Date_time)) %>% 
  filter(!is.na(Date_time)) %>% 
  mutate(Date_time = as_datetime(Date_time)) -> temp_filt_1 # convert back

```

Get an average temperature value per day-time across tanks

```{r}
temp_filt_1 %>% 
  group_by(Date_time) %>% 
  summarise(Mean_temp = mean(Temp_C)) %>% 
  mutate(Day = date(Date_time)) -> temp_time_avgs # create a line for each day-time (can be multiple loggers recording at the same date-time, but want one value as a mean across tanks)
```

# Save averages to a CSV

Here we'll save the full temperature time series, averaged across hobo loggers (1 average value per date-time, NOT the mean daily value)

```{r}
write_csv(x = temp_time_avgs, file = here("generated_data/hobo_timeseries.csv"))
```


